### 2017-11-20 daniel_thesis/ddd_fem.tex
Lit review disguised as documentation and chapters of the thesis.

		Node selection remains an expensive operation and minimising array indexing is of the utmost importance for the best perfomance. Selecting nodes in the traditional sense, i.e. with code branching such as \texttt{if statements} or \texttt{case selection} is unmantainable, verbose and very prone to mistakes. The issue was solved by introducing an auxiliary matrix which defines various parameters that aid node selection and greatly reduces code size, improves readability, and eliminates the need for code branching. The matrix can be constructed utilising \cref{f:fem_node_arr} in order to know which nodes correspond to which \kwd{fe} planes. The $ p\textsuperscript{th} $ column of the matrix\footnote{MATLAB uses column-major ordering, so this gives us the best performance for vectorised code.} corresponds to the $ p\textsuperscript{th} $ plane (according to an arbitrary plane numbering) and is defined as the following column vector,
		\begin{align}\label{e:surf_node_util_vec}
			\vec{V_{p}}^{\mathsf{T}} =
				\begin{bmatrix}
					L_{1p} & L_{2p} & \cdots & L_{Np} & A_{p} & C_{p}
				\end{bmatrix}\,,
		\end{align}	
		where $ L_{np} \equiv $ the numeric label for node $ n $ as given by \cref{f:fem_node_arr}, $ A_{p} \equiv $ the area of the plane, and $ C_{p} \equiv$ the numeric label of the orthogonal coordinate to the plane $ C_{p} = 1, ~2, ~3 $ for the $ x, ~y, ~z $ coordinates respectively. $ A_{p} $ lets us segment our output and transitional arrays so that the only data being modified is that which corresponds to the correct plane and $ C_{p} $ lets us know which coordinate we must use in our selection criteria. Using our particular node labelling scheme (with dimensions $ \Delta x,~ \Delta y,~ \Delta z $ respectively in the $ x,~ y,~ z $ directions), the matrix is defined as
		\begin{align}\label{eq:surf_node_util}
			\mtx{V} = 
				\begin{bmatrix}
					5 & 2 & 6 & 1 & 5 & 4\\
					1 & 6 & 5 & 2 & 6 & 3\\
					8 & 3 & 7 & 4 & 1 & 8\\
					4 & 7 & 8 & 3 & 2 & 7\\
					\Delta y \Delta z & \Delta y \Delta z & \Delta x \Delta z & \Delta x \Delta z & \Delta x \Delta y & \Delta x \Delta y\\
					1 & 1 & 2 & 2 & 3 & 3
				\end{bmatrix}\,.
		\end{align}
		The information codified in \cref{eq:surf_node_util} lets us index and process only the necessary columns to extract the surface nodes we're interested in. The advantage of this setup over a naïve implementation is that it can be relatively easily expanded, maintained, and is general enough that it lends itself to a variety of selection criteria. The columns from left to right (1 to 6) represent: 
		\begin{inparaenum}[{face} 1 $ \equiv $]
			\item $ \min(x),~ yz $-plane;
			\item $ \max(x),~ yz $-plane;
			\item $ \min(y),~ xz $-plane;
			\item $ \max(y),~ xz $-plane;
			\item $ \min(z),~ xy $-plane;
			\item $ \max(z),~ xy $-plane.
		\end{inparaenum}

		It is important to make sure the nodes are extracted in a self consistent manner because the problem definition in \cref{c:lin_rect,c:quad_triang} assumes a specific node ordering. In particular, the labelling chirality must remain the same. This is due to the fact that the problems in \cref{c:lin_rect,c:quad_triang} utilise internal coordinates derived from a set of basis vectors, and scalar projections onto them. If the relationship between the vectors is kept, but the chirality is different (opposite) the calculated quantities will have the wrong sign. If however the relationship between the vectors is different to the problem formulation, the program will crash due to division by 0. Self-consistency was achieved by placing an imaginary observer inside the \kwd{fe} model facing the $ \min(x),~ yz $-plane (face 1) and rotating it to view all 6 planes according to \cref{f:surf_node_plane}.
		\begin{figure}
			\centering
			\includegraphics[width=0.75\linewidth]{surf_node_plane.pdf}
			\caption[Self-consistent, chirality preserving surface planes.]{Self-consistent, chirality preserving surface planes. Plane and node numbering according to the columns of \cref{eq:surf_node_util} and \cref{f:fem_node_arr}.}
			\label{f:surf_node_plane}
		\end{figure}

>>>>

		Node selection remains an expensive operation and minimising array indexing is of the utmost importance for the best perfomance. Selecting nodes in the traditional sense, i.e. with code branching such as \texttt{if statements} or \texttt{case selection} is unmantainable, verbose and very prone to mistakes. The issue was solved by introducing an auxiliary matrix which defines various parameters that aid node selection and greatly reduces code size, improves readability, and eliminates the need for code branching. The matrix can be constructed utilising \cref{f:fem_node_arr} in order to know which nodes correspond to which \kwd{fe} planes. The $ p\textsuperscript{th} $ column of the matrix\footnote{MATLAB uses column-major ordering, so this gives us the best performance for vectorised code.} corresponds to the $ p\textsuperscript{th} $ plane (according to an arbitrary plane numbering) and is defined as,
		\begin{align}\label{e:surf_node_util_vec}
			\vec{V_{p}}^{\mathsf{T}} =
				\begin{bmatrix}
					L_{1p} & L_{2p} & \cdots & L_{Np} & A_{p} & C_{p}
				\end{bmatrix}\,,
		\end{align}	
		where $ L_{np} $ is the numeric label for node $ n $ as given by \cref{f:fem_node_arr}, $ A_{p} $ is the area of the plane, and $ C_{p} $ is the numeric label of the orthogonal coordinate to the plane $ C_{p} = 1, ~2, ~3 $ for the $ x, ~y, ~z $ coordinates respectively. $ A_{p} $ lets us segment our output and transitional arrays so that the only data being modified is that which corresponds to the correct plane and $ C_{p} $ lets us know which coordinate we must use in our selection criteria. Using our particular node labelling scheme (with dimensions $ \Delta x,~ \Delta y,~ \Delta z $ respectively in the $ x,~ y,~ z $ directions), the matrix is defined as,
		\begin{align}\label{e:surf_node_util}
			\mtx{V} = 
				\begin{bmatrix}
					5 & 2 & 6 & 1 & 5 & 4\\
					1 & 6 & 5 & 2 & 6 & 3\\
					8 & 3 & 7 & 4 & 1 & 8\\
					4 & 7 & 8 & 3 & 2 & 7\\
					\Delta y \Delta z & \Delta y \Delta z & \Delta x \Delta z & \Delta x \Delta z & \Delta x \Delta y & \Delta x \Delta y\\
					1 & 1 & 2 & 2 & 3 & 3
				\end{bmatrix}\,.
		\end{align}
		The information codified in \cref{e:surf_node_util} lets us index and process only the necessary columns to extract the surface nodes we're interested in. The advantage of this setup over a naïve implementation is that it can be relatively easily expanded, maintained, and is general enough that it lends itself to a variety of selection criteria. The columns from left to right (1 to 6) represent: 
		\begin{inparaenum}[{face} 1 $ \equiv $]
			\item $ \min(x),~ yz $-plane;
			\item $ \max(x),~ yz $-plane;
			\item $ \min(y),~ xz $-plane;
			\item $ \max(y),~ xz $-plane;
			\item $ \min(z),~ xy $-plane;
			\item $ \max(z),~ xy $-plane.
		\end{inparaenum}

		It is important to make sure the nodes are extracted in a self consistent manner because the problem definitions in \cref{c:lin_rect,c:quad_triang} assume a specific node ordering. In particular, the labelling chirality must remain the same. This is due to the fact that the problems in \cref{c:lin_rect,c:quad_triang} utilise internal coordinates derived from a set of basis vectors, and scalar projections onto them. If the relationship between the vectors is kept, but the chirality is different (opposite) the calculated quantities will have the wrong sign. If however the relationship between the vectors is different to the problem formulation, the program will crash due to division by 0. Self-consistency was achieved by placing an imaginary observer inside the \kwd{fe} model facing the $ \min(x),~ yz $-plane (face 1) and rotating it to view all 6 planes according to \cref{f:surf_node_plane}.
		\begin{figure}
			\centering
			\includegraphics[width=0.75\linewidth]{surf_node_plane.pdf}
			\caption[Self-consistent, chirality preserving surface planes.]{Self-consistent, chirality preserving surface planes. Plane and node numbering according to the columns of \cref{e:surf_node_util} and \cref{f:fem_node_arr}.}
			\label{f:surf_node_plane}
		\end{figure}

### 2017-11-20 daniel_thesis/force_lin_rect_se.tex
Lit review disguised as documentation and chapters of the thesis.

\chapter{Analytical Forces Induced by Dislocations on Linear Rectangular Surface Elements}
\label{c:lin_rect}
	%
	\section{Forces Exerted by a Dislocation Line Segment on Linear Rectangular Surface Elements}
	\label{s:f_lin_rect}
		%
		\begin{figure}
			\centering
			\includegraphics[width=\linewidth]{force_calc_linear_rectangle.pdf}
			\caption[Diagram of the analytical force calculation on linear rectangular surface elements.]{Diagram of the line integral method used to find analytical expressions for the forces exerted by \kwdp{dl} on linear rectangular \kwdp{se} \cite{analytical_integration_of_the_forces_induced_by_dislocations_on_a_surface_element}.
				\textit{1}) For any given point $ \vec{x} $ on the \kwd{se} and any given point $ \vec{x'}$ on the \kwd{dls}, define distance $ R_{a} $.
				\textit{2}) Integrate from $ x_{1} \to x_{2} $ along line direction $ \vec{t} $.
				\textit{3}) Integrate from $ r_{1} \to r_{2} $ along vector $ \vec{p} $.
				\textit{4}) Integrate from $ s_{1} \to s_{2} $ along vector $ \vec{q} $.}
			\label{f:flrs}
		\end{figure}
		\subsection{Resolving Singularities when Dislocation Line Segments are Parallel to Surface Elements}
		\label{ss:par_dln_se}
			%
			\begin{figure}
				\centering
				\includegraphics[width=\linewidth]{ftot_rotation_lin_rect.pdf}
				\caption[Avoiding singularities by rotating dislocation line segments.]{Effects of rotating a single \kwd{dls} on the forces exerted by it on a linear rectangular \kwd{se}. The specific values of this function are not known \emph{a priori}, all that is known is that it must be periodic ($ T = 2\pi$) and have finite maximum and minimum values. The singularity is avoided by perturbing the angle $ \theta = 0 \to \theta = \pm \epsilon\,\, \epsilon \gtrsim 0 $.}
				\label{f:rflrs}
			\end{figure}
		%
\savearabiccounter

>>>>

\chapter{Analytical Forces Induced by Dislocations on Linear Rectangular Surface Elements}
\label{c:lin_rect}
	%
	\section{Forces Exerted by a Dislocation Line Segment on Linear Rectangular Surface Elements}
	\label{s:f_lin_rect}
		%
		Coupling \kwd{ddd} to \kwd{fem} requires the traction field $ \tns{\sigma^{\infty}}(\vec{x}) \cdot \vec{n}(\vec{x}) $ to be distributed among the set of relevant discrete nodes of a \kwd{fe} or \kwd{be} model. This is usually achieved as follows,
		\begin{align}\label{e:ddd_fem_force}
			\vec{F}^{(n)} = \int_{S^{e}} \left[\tns{\sigma^{\infty}}(\vec{x}) \cdot \vec{n}(\vec{x})\right] L_{n}(\vec{x})\; \rvar{d}S_{e}\,,
		\end{align}
		where $ \rvar{d}S_{e} $ is the infinitesimal surface element with surface area $ S_{e} $. $ L_{n}(\vec{x}) $ are so-called shape functions (interpolation functions) that distribute  the traction field among the surface element's nodes.

		The problematic singularity associated with the classical Volterra dislocation is avoided by using the non-singular formulation of \citet{a_non-singular_continuum_theory_of_dislocations}. The stress field of a dislocation in a homogenous infinite linear elastic domain can be calculated as a contour integral along the loop \cite{mura_t},
		\begin{align}\label{e:field_stress}
			\sigma_{ij}^{\infty}(\vec{x}) = C_{ijkl} \oint \epsilon_{lnh} C_{pqmn} \dfrac{\partial G_{kp}(\vec{x} - \vec{x'})}{\partial x_{q}} b_{m} \rvar{d}x'_{h}\,,
		\end{align}
		where $ C_{ijkl} $ is the elastic stiffness matrix, $ \epsilon_{lnh} $ the permutation operator, \kwds{bv} the \kwd{bv}, $ \vec{x'} $ the coordinate that spans the dislocation, and $ G_{kp}(\vec{x} - \vec{x'}) $ is Green's function of elasticity \cite{mura_t}. $ G_{kp}(\vec{x} - \vec{x'}) $ is defined as the displacement component in the $ x_{k} $ direction at point $ \vec{x} $ due to a force applied in the $ x_{p} $ direction at point $ \vec{x'} $. The traditional singularity comes from taking the \kwd{bv} distribution as a delta function. \citet{a_non-singular_continuum_theory_of_dislocations} proposed an alternative definition of $ G_{kp}(\vec{x} - \vec{x'}) $ which has a wider isotropic spread mainly localised in a radius $ a $ around the dislocation core,
		\begin{align}\label{e:elastic_green_func}
			G_{ij}(\vec{x} - \vec{x'}) = \dfrac{1}{8\pi \mu}\left[ \delta_{ij} \partial_{pp} - \dfrac{1}{2(1-\nu)} \partial_{ij} \right] R_{a}\,,
		\end{align}
		where $ \mu $, $ \nu $ are the isotropic shear modulus and Poisson's ratio respectively, $ \delta_{ij} $ is the Kronecker Delta, $ \partial_{x_{1} \ldots\, x_{n}} \equiv \dfrac{\partial^{n}}{\partial x_{1} \ldots\, \partial x_{n}}$. $ R_{a} $ is the defined as,
		\begin{subequations}
			\begin{align}\label{e:ra}
				R_{a}(\vec{x}) &= R(\vec{x}) * w(\vec{x}) = \int R(\vec{x} - \vec{x'}) w(\vec{x'}) \rvar{d}^{3}\vec{x'}\nn
							   &= \sqrt{R^{2} + a^{2}}\\
				w(\vec{x}) 	   &= \dfrac{15 a^{4}}{8\pi (R^{2} + a^{2})^{7/2}}\,,
			\end{align}
		\end{subequations}
		 where $ w(\vec{x}) $ is a \kwd{bv} distribution in \cite{a_non-singular_continuum_theory_of_dislocations}. To keep in line with the definition of \cref{e:field_stress}, where the displacement component is in the $ x_{k} $ direction, $ R_{a} = \sqrt{R_{k}^{2} + a^{2}} $.

		 For two dislocation nodes (1, 2) connected by straight line segments \cref{e:field_stress} becomes,
		 \begin{align}
			 \tns{\sigma}^{(12)}(\vec{x}) =
			 \begin{split} 
				&-\dfrac{\mu}{8\pi} \int\limits_{\vec{x}_{1}}^{\vec{x}_{2}} \left( \dfrac{2}{R_{a}^{3}} + \dfrac{3 a^{2}}{R_{a}^{5}} \right) \left[ \left( \vec{R} \times \vec{b} \right) \otimes \rvar{d}\vec{x'} + \rvar{d}\vec{x'} \otimes \left( \vec{R} \times \vec{b} \right) \right]\\
				%
				&+ \dfrac{\mu}{4\pi (1 - \nu)} \int\limits_{\vec{x}_{1}}^{\vec{x}_{2}} \left( \dfrac{1}{R_{a}^{3}} + \dfrac{3 a^{2}}{R_{a}^{5}} \right) \left[ \left(\vec{R} \times \vec{b}\right)\cdot \rvar{d}\vec{x'} \right] \mtx{I}_{2}\\
				%
				&- \dfrac{\mu}{4\pi (1 - \nu)} \int\limits_{\vec{x}_{1}}^{\vec{x}_{2}} \dfrac{1}{R_{a}^{3}} \left[ \left(\vec{b} \times \rvar{d}\vec{x'}\right) \otimes \vec{R} + \vec{R} \otimes \left(\vec{b} \times \rvar{d}\vec{x'}\right) \right]\\
				&+ \dfrac{\mu}{4\pi (1 - \nu)} \int\limits_{\vec{x}_{1}}^{\vec{x}_{2}} \dfrac{3}{R_{a}^{5}} \left[ \left( \vec{R} \times \vec{b} \right) \cdot \rvar{d}\vec{x'} \right]\vec{R} \otimes \vec{R}
			 \end{split}
		 \end{align}
		\begin{figure}
			\centering
			\includegraphics[width=\linewidth]{force_calc_linear_rectangle.pdf}
			\caption[Diagram of the analytical force calculation on linear rectangular surface elements.]{Diagram of the line integral method used to find analytical expressions for the forces exerted by \kwdp{dl} on linear rectangular \kwdp{se} \cite{analytical_integration_of_the_forces_induced_by_dislocations_on_a_surface_element}.
				\textit{1}) For any given point $ \vec{x} $ on the \kwd{se} and any given point $ \vec{x'}$ on the \kwd{dls}, define distance $ R_{a} $.
				\textit{2}) Integrate from $ x_{1} \to x_{2} $ along line direction $ \vec{t} $.
				\textit{3}) Integrate from $ r_{1} \to r_{2} $ along vector $ \vec{p} $.
				\textit{4}) Integrate from $ s_{1} \to s_{2} $ along vector $ \vec{q} $.}
			\label{f:flrs}
		\end{figure}
		\subsection{Resolving Singularities when Dislocation Line Segments are Parallel to Surface Elements}
		\label{ss:par_dln_se}
			%
			\begin{figure}
				\centering
				\includegraphics[width=\linewidth]{ftot_rotation_lin_rect.pdf}
				\caption[Avoiding singularities by rotating dislocation line segments.]{Effects of rotating a single \kwd{dls} on the forces exerted by it on a linear rectangular \kwd{se}. The specific values of this function are not known \emph{a priori}, all that is known is that it must be periodic ($ T = 2\pi$) and have finite maximum and minimum values. The singularity is avoided by perturbing the angle $ \theta = 0 \to \theta = \pm \epsilon\,\, \epsilon \gtrsim 0 $.}
				\label{f:rflrs}
			\end{figure}
		%
\savearabiccounter

### 2017-11-20 daniel_thesis/para_force_se.tex
Lit review disguised as documentation and chapters of the thesis.

	where $ e\equiv $~\kwd{se} label, $ n \equiv $~node label, $ E \equiv $~number of \kwdp{se} in the scope. $ N \equiv $~number of nodes in a \kwd{se}.

@@

	where $ e\equiv $~surface \textbf{e}lement, $ n \equiv $~\textbf{n}ode, $ E \equiv $~total number of \kwdp{se} in scope, $ N \equiv $~total number of nodes in each \kwd{se}.

@@

	\subsection{Test Case}
	\begin{figure}
		\includegraphics[width=\linewidth]{cuda_test.pdf}
		\caption[Test case for CUDA development.]{Test case for CUDA development. Dislocation line not shown in node separation as the parallelisation happens across surface elements. Not to perspective or scale.}
		\label{f:cuda_test}
	\end{figure}
	According to \cref{a:fcne} and the node labelling scheme of \cref{f:flrs} the data mapping would look like:
	\begin{align}\label{e:cuda_test}
		\vec{X}^{\textrm{SE}} = 
		\begin{split}
			\left[\overbrace{\underbrace{1,1,0,0}_{\mathbf{x_{3}}}, \underbrace{2,2,1,1}_{\mathbf{x_{4}}},
				\underbrace{1,1,0,0}_{\mathbf{x_{5}}}, \underbrace{2,2,1,1}_{\mathbf{x_{6}}}}^{x\textrm{-coord}},\right.\\
			\overbrace{\underbrace{0,1,1,0}_{\mathbf{x_{3}}}, \underbrace{0,1,1,0}_{\mathbf{x_{4}}},
				\underbrace{1,2,2,1}_{\mathbf{x_{5}}}, \underbrace{1,2,2,1}_{\mathbf{x_{6}}}}^{y\textrm{-coord}},\\
			\left.\overbrace{\underbrace{0,0,0,0}_{\mathbf{x_{3}}}, \underbrace{0,0,0,0}_{\mathbf{x_{4}}},
				\underbrace{0,0,0,0}_{\mathbf{x_{5}}}, \underbrace{0,0,0,0}_{\mathbf{x_{6}}}}^{z\textrm{-coord}}\right]
		\end{split}
	\end{align}
	In general the number of indices in the array would be $ 3 \times e \times n $ where $ e \equiv $~number of surface elements, $ n \equiv $~number of nodes per element and 3 is the number of spatial dimensions.
	%
	\section{Thread Block Size Optimisation}
	\href{http://docs.nvidia.com/cuda/cuda-c-best-practices-guide/index.html#memory-optimizations}{http://docs.nvidia.com/cuda/cuda-c-best-practices-guide/index.html\#memory-optimizations}
	\href{http://docs.nvidia.com/cuda/cuda-c-best-practices-guide/index.html#execution-configuration-optimizations}{http://docs.nvidia.com/cuda/cuda-c-best-practices-guide/index.html\#execution-configuration-optimizations}
	\savearabiccounter

>>>>

	where $ e $ is the \kwd{se} label, $ n $ the node label, $ E $ the number of \kwdp{se} in the scope. $ N $ the number of nodes in a \kwd{se}.

@@

	where $ e $ is the surface \textbf{e}lement, $ n $ the \textbf{n}ode, $ E $ the total number of \kwdp{se} in scope, $ N $ the total number of nodes in each \kwd{se}.

@@

	\subsection{Test Case}
	\begin{figure}
		\includegraphics[width=\linewidth]{cuda_test.pdf}
		\caption[Test case for CUDA development.]{Test case for CUDA development. Dislocation line not shown in node separation as the parallelisation happens across surface elements. Not to perspective or scale.}
		\label{f:cuda_test}
	\end{figure}
	According to \cref{a:fcne} and the node labelling scheme of \cref{f:flrs} the data mapping would look like:
	\begin{align}\label{e:cuda_test}
		\vec{X}^{\textrm{SE}} = 
		\begin{split}
			\left[\overbrace{\underbrace{1,1,0,0}_{\mathbf{x_{3}}}, \underbrace{2,2,1,1}_{\mathbf{x_{4}}},
				\underbrace{1,1,0,0}_{\mathbf{x_{5}}}, \underbrace{2,2,1,1}_{\mathbf{x_{6}}}}^{x\textrm{-coord}},\right.\\
			\overbrace{\underbrace{0,1,1,0}_{\mathbf{x_{3}}}, \underbrace{0,1,1,0}_{\mathbf{x_{4}}},
				\underbrace{1,2,2,1}_{\mathbf{x_{5}}}, \underbrace{1,2,2,1}_{\mathbf{x_{6}}}}^{y\textrm{-coord}},\\
			\left.\overbrace{\underbrace{0,0,0,0}_{\mathbf{x_{3}}}, \underbrace{0,0,0,0}_{\mathbf{x_{4}}},
				\underbrace{0,0,0,0}_{\mathbf{x_{5}}}, \underbrace{0,0,0,0}_{\mathbf{x_{6}}}}^{z\textrm{-coord}}\right]
		\end{split}
	\end{align}
	In general the number of indices in the array would be $ 3 \times e \times n $ where $ e $ is the number of surface elements, $ n $ the number of nodes per element and 3 the number of spatial dimensions.
	%
	\section{Thread Block Size Optimisation}
	\href{http://docs.nvidia.com/cuda/cuda-c-best-practices-guide/index.html#memory-optimizations}{http://docs.nvidia.com/cuda/cuda-c-best-practices-guide/index.html\#memory-optimizations}
	\href{http://docs.nvidia.com/cuda/cuda-c-best-practices-guide/index.html#execution-configuration-optimizations}{http://docs.nvidia.com/cuda/cuda-c-best-practices-guide/index.html\#execution-configuration-optimizations}
	\savearabiccounter
